---
title: "Survival"
output:
  html_document:
    df_print: paged
 
---

Here we describe the parameters associated with immature and mature survival.

**Table 1**: Summary of model parameters associated with survival.


| In-text | In-code |    Definition        | 
|:--------:|:--------:|:-------------------------------------|
|$\color{red}{a_{i,i}}$ |$\color{red}{aii.vec}$ | Survival probability of an immature individual for each cytotype.  	|
|$\color{red}{\mu}$ | $\color{red}{as.msurv}$	| Survival probability of an mature individual for each cytotype. 	|
|$\color{red}{ci}$| $\color{red}{env.ci}$		| Proportion of environmental variance used to define mature survival rate per generation.|
                                                                

## Immature survival  

Immature probability of survival was defined as followed:
     
$$S_{i, 1}(t) = e^{-aii[j]*\sum_{j=1}^{2}c_{i,j}(t)}$$
Where $c_{i, 1}(t + 1)$ is defined by a binominal distribution with $c_{i,1}(t)$ trials and probability of success equal to $S_{i,1}(t)$. As aii increases, the number of individuals expected to survive decreases (**Figure 1**).   


```{r immature, echo=FALSE, fig.align='center', fig.height = 10, fig.width = 8}
poploop <- seq(10, 100000, by = 50)


is1 <- 0.00001
is1out <- c()
is2 <- 0.00005
is2out <- c()
is3 <- 0.0001
is3out <- c()
is4 <- 0.0005
is4out <- c()
is5 <- 0.001
is5out <- c()

for(i in 1:200){
  is1out[i] <- rbinom(n = 1, size = poploop[i], prob = exp(-is1*poploop[i]))
  is2out[i] <- rbinom(n = 1, size = poploop[i], prob = exp(-is2*poploop[i]))
  is3out[i] <- rbinom(n = 1, size = poploop[i], prob = exp(-is3*poploop[i]))
  is4out[i] <- rbinom(n = 1, size = poploop[i], prob = exp(-is4*poploop[i]))
  is5out[i] <- rbinom(n = 1, size = poploop[i], prob = exp(-is5*poploop[i]))
}

is1_plot <- data.frame(popsize = poploop, out = is1out)
is2_plot <- data.frame(popsize = poploop, out = is2out)
is3_plot <- data.frame(popsize = poploop, out = is3out)
is4_plot <- data.frame(popsize = poploop, out = is4out)
is5_plot <- data.frame(popsize = poploop, out = is5out)

par(mfrow = c(5, 1), mar=c(2,2,2,2))
plot(x = is1_plot$popsize, y = is1_plot$out, col = "red", pch = 16, type = "b",
     xlab = "Population Size",  ylab = "survival", main="aii = 0.00001", xlim = c(0, 100050), ylim = c(0, 10000))
plot(x = is2_plot$popsize, y = is2_plot$out, col = "red", pch = 16, type = "b",
     xlab = "Population Size",  ylab = "survival", main="aii = 0.00005", xlim = c(0, 100050), ylim = c(0, 10000))
plot(x = is3_plot$popsize, y = is3_plot$out, col = "red", pch = 16, type = "b",
     xlab = "Population Size",  ylab = "survival", main="aii = 0.0001", xlim = c(0, 100050), ylim = c(0, 10000))
plot(x = is4_plot$popsize, y = is4_plot$out, col = "red", pch = 16, type = "b",
     xlab = "Population Size",  ylab = "survival", main="aii = 0.0005", xlim = c(0, 100050), ylim = c(0, 10000))
plot(x = is5_plot$popsize, y = is5_plot$out, col = "red", pch = 16, type = "b",
     xlab = "Population Size",  ylab = "survival", main="aii = 0.001", xlim = c(0, 100050), ylim = c(0, 10000))


```
**Figure 1**: Expected number of individuals to survive given different immature survival probabilities.

<br>

## Mature survival   
Mature probability of survival was defined using a beta distribution to sample mature survival probability at each generation based on the mean mature survival probability ($\mu$) and  proportion of variance ($env.ci$). Where $c_{i, 2}(t + 1)$ is defined by based on a binominal distribution with $c_{i,2}(t)$ trials and probability of success equal to $S_{i,2}(t)$. The survival probability, $S_{i,2}(t)$, varies over time and is equal to $Beta(\alpha, \beta)$. 

We can define $\alpha$ and $\beta$ based on $\mu$ and $env.ci$ by first defining variance as $\sigma^{2} = ci*\mu(1-\mu)$, then $\alpha = \mu * \frac{\mu*(1-\mu)}{(\sigma^{2} - 1)}$ and $\beta = (1-\mu)*\frac{\mu*(1-\mu)}{(\sigma^{2} - 1)}$. 

Note,  $0 \le env.ci < 1$, see [Gaynor et al. 2023](https://doi.org/10.1101/2023.03.29.534764) for details. As env.ci increases, we see the variance in probability of survival also increase (**Figure 2**). Additionally, as the mature survival increase, so does the sampled probability of survival.
  

```{r mature, echo=FALSE, fig.align='center', fig.height = 8, fig.width = 8}
source("../R/mature.surv.calc.R")
source("../R/var.option.R")
source("../R/alphabeta.calc.R")
envi.ci.list <- seq(0.05, 0.95, by = 0.05) 
as.msurv.list <- seq(0.05, 0.95, by = 0.05)

out.set.me <-  matrix(nrow = (19*21), ncol = 3)
klist <- list()

out.set.list <- list()
for(k in 1:100){
  out.set.list[[k]] <- matrix(nrow = (19*19), ncol = 4)
  c <- 0
for(i in 1:19){
  env.ci <- envi.ci.list[i]
  for(j in 1:19){
    c <- c + 1
    out.set.list[[k]][c, 1] <- env.ci[1]
    as.msurv <- as.msurv.list[j]
    out.set.list[[k]][c, 2] <- as.msurv
    out.set.list[[k]][c, 3] <-  mature.surv.calc(env.ci = env.ci, as.msurv = as.msurv)

  }
}
}
out.set.me <- do.call(rbind, out.set.list)
out.set.me.df <- as.data.frame(out.set.me)

library(ggplot2)
ggplot(out.set.me.df, aes(x = V2, y = V3, col = V1)) +
  geom_point() +
  #geom_line() +
  xlab("Mature Survival Probability") +
  ylab("Sampled Probability of Survival") +
  facet_wrap(~V1)
  

```
**Figure 2**: Expected survival probability when we sample a beta distribution based on the mean mature survival probability ($\mu$) (x-axis) and  proportion of variance ($env.ci$) (plot title).


