---
title: "Survival"
output:
  html_document:
    df_print: paged
---

Here we describe the parameters associated with immature and mature survival.

**Table 1**: Summary of model parameters associated with survival.


| In-text | In-code |    Definition        | 
|:--------:|:--------:|:-------------------------------------|
|$\color{red}{\rho}$ |$\color{red}{rho}$ | Rank correlation between surivival of different cytotypes.|
|$\color{red}{a_{i}}$ |$\color{red}{aii.vec}$ | Cytotype-specific response to competition related to immature survival.|
|$\color{red}{\alpha_{i,k}}$|$\color{red}{is.density.type}$ | Type-specific density dependence. Competitive effect of cytotype $k$ on cytotype $i$'s survival. Options: "all" or "like-cytotype". When equal to "all", $\alpha_{i,k} = 1$. Similarly, when equal to "like-cytotype", $\alpha_{i,k} = 0$ | 
|$\color{red}{y_{j}}$| $\color{red}{yj}$| Stage-specific density dependence. |
||$\color{red}{env.immature.survival}$| Allows environmental stochasticity to influences immature survival. Defaults as "TRUE"|
|$\color{red}{\sigma}$|$\color{red}{env.sigma}$| Proportion of environmental variance for immature individual. |
|$\color{red}{\mu}$ | $\color{red}{as.msurv}$	| Survival probability of an mature individual for each cytotype. 	|
|$\color{red}{ci}$| $\color{red}{env.ci}$		| Proportion of environmental variance used to define mature survival rate per generation.	|
                                                                

## Immature survival  
Immature survival is defined as:
$$\hat{s}_{i,1}(t) = \frac{e^{\mu_{i,1}}}{1+e^{\mu_{i,1}}} = \text{exp}\left\{-a_i\sum_{k=2}^{4}\alpha_{i,k} \sum_{j=1}^{2}y_jc_{i,k}(t)\right\}$$

where $\mu_{i,1}$ is the logs-odd survival, ${y_{j}}$ is the stage-specific density dependency, $a_i$ is the cytotype-specific response to competition related to immature survival, $\alpha_{i,k}$ is type-specific density dependence, and $c_{i,k}$ is the number of individuals of cytotype k (k = \{2, 3, 4\}) and stage j (j = = \{1,2\}).

As aii increases, the probability of survival decreases at a lower population size (**Figure 1**). When $\sigma$ increases, environmental variance increases.  Therefore, $\sigma$ (env.sigma) increases, the propability of survival becomes more variable. 



```{r immature, echo=FALSE, fig.align='center', fig.height = 15, fig.width = 8}
library(ggplot2)
source("../R/env.copula.R")
Xi.is <- (env.copula(0, 2000))[[1]]
imsurv <- function(isval, env.sigma){
   isout <- c() 
   for(i in 1:2000){
    immature.det.surv <-  exp(-isval*poploop[i])
    immature.mu <- log(immature.det.surv/(1-immature.det.surv))
     as.isurv <- (exp(immature.mu + (env.sigma*Xi.is[i]))/(1 + exp(immature.mu + (env.sigma*Xi.is[i]))))
      isout[i] <- as.isurv
   #isout[i] <-  rbinom(n = 1, size = poploop[i], prob = as.isurv)
}
      return(data.frame(popsize = poploop, out = as.numeric(isout), env.sigma =  env.sigma, val = paste0("aii = ", isval)))
}

poploop <- seq(10, 100000, by = 50)


is1 <- 0.00001
is1out <- imsurv(is1, 1)
is1outb <- imsurv(is1, 25)
is1outc <- imsurv(is1, 50)
is1outall <- rbind(is1out, is1outb, is1outc)
is2 <- 0.00005
is2out <- imsurv(is2, 1)
is2outb <- imsurv(is2, 25)
is2outc <- imsurv(is2, 50)
is3 <- 0.0001
is3out <- imsurv(is3, 1)
is3outb <- imsurv(is3, 25)
is3outc <- imsurv(is3, 50)
is4 <- 0.0005
is4out <- imsurv(is4, 1)
is4outb <- imsurv(is4, 25)
is4outc <- imsurv(is4, 50)
is5 <- 0.001
is5out <- imsurv(is5, 1)
is5outb <- imsurv(is5, 25)
is5outc <- imsurv(is5, 50)
is6 <- 0.005
is6out <- imsurv(is6, 1)
is6outb <- imsurv(is6, 25)
is6outc <- imsurv(is6, 50)

isoutall <- rbind(is1out, is1outb, is1outc, 
                   is2out, is2outb,is2outc, 
                   is3out, is3outb, is3outc,
                   is4out, is4outb, is4outc,
                   is5out, is5outb, is5outc, 
                  is6out, is6outb, is6outc)
isoutall$val <- factor(isoutall$val, levels =c("aii = 0.001", "aii = 0.005", "aii = 0.0001", "aii = 0.0005", "aii = 0.00001", "aii = 0.00005"))
options(scipen=999)

ggplot(isoutall, aes(x = popsize, y = out, col = factor(env.sigma)))+
  geom_point(alpha = 0.9) +
  scale_color_manual(values = c("blue", "purple", "red")) +
  facet_wrap(factor(val) ~ factor(env.sigma), nrow = 6, ncol = 3) +
  labs(color = "env.sigma") +
  xlab("Population Size") +
  ylab("Survival Probability")


```
**Figure 1**: Probability of survival as a function of population size and environmental variance. 

<br>

## Mature survival   
Mature probability of survival was defined using a beta distribution to sample mature survival probability at each generation based on the mean mature survival probability ($\mu$) and  proportion of variance ($env.ci$). Where $c_{i, 2}(t + 1)$ is defined by based on a binominal distribution with $c_{i,2}(t)$ trials and probability of success equal to $S_{i,2}(t)$. The survival probability, $S_{i,2}(t)$, varies over time and is equal to $Beta(\alpha, \beta)$. 

We can define $\alpha$ and $\beta$ based on $\mu$ and $env.ci$ by first defining variance as $\sigma^{2} = ci*\mu(1-\mu)$, then $\alpha = \mu * \frac{\mu*(1-\mu)}{(\sigma^{2} - 1)}$ and $\beta = (1-\mu)*\frac{\mu*(1-\mu)}{(\sigma^{2} - 1)}$. 

Note,  $0 \le env.ci < 1$, see [Gaynor et al. 2025](https://doi.org/10.1086/734411) for details. As env.ci increases, we see the variance in probability of survival also increase (**Figure 2**). Additionally, as the mature survival increase, so does the sampled probability of survival.
  

```{r mature, echo=FALSE, fig.align='center', fig.height = 8, fig.width = 8}
source("../R/mature.surv.calc.R")
source("../R/var.option.R")
source("../R/alphabeta.calc.R")
source("../R/surv.shape.R")
envi.ci.list <- seq(0.05, 0.95, by = 0.05) 
as.msurv.list <- seq(0.05, 0.95, by = 0.05)

out.set.me <-  matrix(nrow = (19*21), ncol = 3)
klist <- list()
Ui.vec <- (env.copula(0, 2000))[[2]]

out.set.list <- list()
for(k in 1:100){
  out.set.list[[k]] <- matrix(nrow = (19*19), ncol = 4)
  c <- 0
for(i in 1:19){
  env.ci <- envi.ci.list[i]
  for(j in 1:19){
    c <- c + 1
    out.set.list[[k]][c, 1] <- env.ci[1]
    as.msurv <- as.msurv.list[j]
    out.set.list[[k]][c, 2] <- as.msurv
    msurv.alpha.beta <- surv.shape(env.ci = env.ci, raw.means = as.msurv)
           out.set.list[[k]][c, 3] <- qbeta(Ui.vec[k,1], msurv.alpha.beta[1,1], msurv.alpha.beta[2,1])
    #out.set.list[[k]][c, 3] <-  mature.surv.calc(env.ci = env.ci, as.msurv = as.msurv)

  }
}
}
out.set.me <- do.call(rbind, out.set.list)
out.set.me.df <- as.data.frame(out.set.me)

library(ggplot2)
ggplot(out.set.me.df, aes(x = V2, y = V3, col = V1)) +
  geom_point() +
  #geom_line() +
  xlab("Mature Survival Probability") +
  ylab("Probability of Survival") +
  facet_wrap(~V1)





```
**Figure 2**: Expected survival probability when we sample a beta distribution based on the mean mature survival probability ($\mu$) (x-axis) and  proportion of variance ($env.ci$) (plot title).



